---
title: "CCLHDME CCSR Tool Demo"
format: html
editor: visual
---

## CCLHDME CCSR Tool Demo

Load libraries

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
```

Load sample HCAI PDD data, CCSR categories, and CCSR reference details

```{r}
#synthetic data for example
sample_data <- read_xlsx("G:\\Public Health\\LHNC\\Epidemiology Team\\Fellow Projects\\Alejandra Cal-EIS 2025-2026\\CCLHDME Tool Demo\\Sample_HCAI_Data.xlsx")

#download file and sheet that contains the CCSR translations 
CCSR_full <- read_xlsx("G:\\Public Health\\LHNC\\Epidemiology Team\\Fellow Projects\\Alejandra Cal-EIS 2025-2026\\CCLHDME Tool Demo\\CCSR References\\Diagnosis and External Cause Codes\\DXCCSR-Reference-File-v2025-1.xlsx", sheet = 4)
```

Fixing column names:

```{r}
################### DO NOT RUN MORE THAN ONCE #################
colnames(CCSR_full) <- CCSR_full[1, ]
CCSR_full <- CCSR_full[-c(1), ]
#you can also just manually delete the first row heading from the excel sheet in the CCSR tool

##renaming the column, can rename other columns here
colnames(CCSR_full)[colnames(CCSR_full) == "ICD-10-CM Code"] <- "ICD_10_Code" 
```

Restricting CCSR to only Inpatient default codes since we are working with hospitalization data

```{r}
clean_CCSR_full <- CCSR_full %>%
  filter(`Inpatient Default CCSR (Y/N/X)`== "Y" | `Inpatient Default CCSR (Y/N/X)`== "X") #x for external cause codes
```

CCSR mapping tool for ICD-10-CM codes to CCSR Categories: Good for identifying body systems but will still need to be translated to the category description file using join for full name of diagnosis

```{r}
#creating the map
ccsr_map <- setNames(clean_CCSR_full$`CCSR Category`, clean_CCSR_full$ICD_10_Code)

#identify which columns need to be translated
colnames(sample_data)

#creating new dataset so that original is unedited
sample_data_ccsr <- sample_data

#applying map to the data
sample_data_ccsr[2:8] <- lapply(sample_data[2:8], function(col){
  as.character(ccsr_map[as.character(col)])
})
```

CCSR mapping tool for ICD-10-CM codes to CCSR Category Description

```{r}
#same thing as above
ccsr_des_map <- setNames(clean_CCSR_full$`CCSR Category Description`, clean_CCSR_full$ICD_10_Code)

sample_data_ccsr_des <- sample_data

sample_data_ccsr_des[2:8] <- lapply(sample_data[2:8], function(col){
  as.character(ccsr_des_map[as.character(col)])
})
```

### EXAMPLES

#### Diagnosis Columns

Looking at primary diagnosis column using CCSR category description

```{r}
pri_diag_freq <- sample_data_ccsr_des %>%
  group_by(diag_p)%>%
  summarize(count= n())
```

Looking through all diagnosis columns

```{r}
#identifying columns that need to be translated
colnames(sample_data_ccsr_des)

#pivoting longer so that each diagnosis can be counted
all_diag_freq <- sample_data_ccsr_des %>%
  pivot_longer(
    cols = c(2:5),
    values_to = "diagnosis"
  ) %>%
  group_by(diagnosis)%>%
  summarize(count= n())
```

#### External Cause Codes

First column:

```{r}
pri_ecm_freq <- sample_data_ccsr_des %>%
  group_by(ecm1)%>%
  summarize(count= n())
```

All columns:

```{r}
colnames(sample_data_ccsr_des)

all_ecm_freq <- sample_data_ccsr_des %>%
  pivot_longer(
    cols = c(6:8),
    values_to = "ext_cause_codes"
  ) %>%
  group_by(ext_cause_codes)%>%
  summarize(count= n())
```

#### Procedure Codes

Load in the CCSR for procedures

```{r}
prccsr <- read_xlsx("G:\\Public Health\\LHNC\\Epidemiology Team\\Fellow Projects\\Alejandra Cal-EIS 2025-2026\\CCLHDME Tool Demo\\CCSR References\\Procedure codes\\PRCCSR-ReferenceFile-v2026-1.xlsx", sheet = 5)
```

Creating the map

```{r}
ccsr_pcs_map <- setNames(prccsr$`CCSR Procedure Category Description`, prccsr$'ICD-10-PCS Code')

sample_data_ccsr_des[9:11] <- lapply(sample_data_ccsr_des[9:11], function(col){
  as.character(ccsr_pcs_map[as.character(col)])
})
```

Getting counts for primary procedure

```{r}
pri_proc_freq <- sample_data_ccsr_des %>%
  group_by(proc_p) %>%
  summarise(counts = n())
```

Getting counts for all procedure fields

```{r}
all_proc_freq <- sample_data_ccsr_des %>%
  pivot_longer(
    cols = c(9:11),
    values_to = "proc_codes"
  ) %>%
  group_by(proc_codes)%>%
  summarize(count= n())
```

```{r}

```
